The hidden manifold distance for functional data

### 2019-08-23 index.tex
.gitignore fix

The clustering and classification of functional data, discretely-sampled
data varying over a continuum, is frequently encountered in machine
learning. While one can naively treat the data as vectors in Euclidean
space, one may also regard each data point as a function. The challenge
of analyzing infinite-dimensional functions can be ameliorated if the
curves are smooth since the dimensionality is then only artificially
high. This insight underpins functional data analysis, a subfield of
statistics that studies functional data. For certain types of functional
data such as classes of probability density functions and classes of
warped curves of a common template function, there is further structure
to exploit. Namely, when functional data lie on a manifold, the
dimensionality at play might be even lower if the manifold is itself of
low intrinsic dimension. This work addresses the estimation of pairwise
geodesic distances between functional manifold data that are observed
with noise and which may live near, rather than on, a manifold. The
proposed methodology first sends the observed functional data to the
hidden manifold, estimated using subspace-constrained mean-shift.
Geodesic distances are subsequently calculated by employing
shortest-path algorithms on this estimated manifold. Improved estimation
of the pairwise geodesic distance has beneficial implications for many
downstream tasks in functional data analysis which we illustrate in the
case of distance-based functional classification.

@@

???Still need to describe parameters under study: SNR, denseness of
observations (\(K_{observed}\)). Describe parameters that are fixed:
\(K_smooth\) and com\_grid. Describe tuning parameters for our method:
\(s\) and \(h\).

\begin{figure}
\includegraphics{sim_results/sce=5_SNR=high_Kobs=30}
\caption{blah}
\end{figure}

>>>>

The clustering and classification of functional data, discretely-sampled
curves varying over a continuum, is frequently encountered in machine
learning. While one can naively treat the data as vectors in Euclidean
space, one may also regard the data as functions. The challenge of
analyzing infinite-dimensional objects can be ameliorated if the curves
are smooth since the dimensionality is then only artificially high. This
insight underpins functional data analysis, a subfield of statistics
that studies functional data. For certain types of functional data such
as probability density functions and warped curves of a common template
function, there is further structure to exploit. Namely, certain
functional data may lie on a manifold of low intrinsic dimension. For
nonlinear manifolds, even basic operations such as addition and
subtraction require special consideration. This work addresses the
estimation of pairwise geodesic distances between noisy functional
manifold data that live near, rather than on, a manifold. This setting
cannot be tackled by classic manifold learning techniques which require
data to live exactly on a manifold. The proposed methodology first sends
the observed functional data to the hidden manifold, estimated using
subspace-constrained mean-shift. Geodesic distances are subsequently
calculated by employing shortest-path algorithms on this estimated
manifold. Improved estimation of the pairwise geodesic distance has
beneficial implications for many downstream tasks in functional data
analysis which we illustrate in the case of distance-based functional
classification/clustering.

@@

The tuning parameters for the proposed method were chosen as described
at the end of Section
\ref{proposed-method-for-estimating-geodesic-distances}. In step 1 of
our procedure, we take \(D = 100\) for the size of the common time grid.
We will investigate how two parameters affect the performance of the
various methods in consideration. The first one is the signal-to-noise
ratio ???describe??? which we fix at either \(0.1\) (low) or \(0.5\)
(high). The second parameter is \(K\), the dimension of the
dscrietly-observed noisy \(Y_i\)'s. We set \(K\) to either \(30\)
(sparse) or \(100\) (abundant).

We show the performance of the SS and L2 method versus our method for
three different values of \(s\). We omit the result of the square root
velocity scenario because all methods perform comparably. We can see
from Figures \ref{fig:normal} and \ref{fig:warping} that ???.

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
\caption{Normal density}
\label{fig:normal}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
\caption{Warping function}
\label{fig:warping}
\end{figure}
